Respiratory-Related Mortality Rates Show A Positive Correlation With Increasing Air pollution

### 2024-03-26 paper/paper.qmd
updates in model

deaths_data <- read.csv(here("data/analysis_data/cleaned_file.csv"))
air_quality_data <- read.csv(here("data/analysis_data/cleaned_air_data.csv"))
data <- read.csv(here("data/analysis_data/cleaned_chart_data.csv"))
data_peak <- read.csv(here("data/analysis_data/cleaned_peak_data.csv"))
data_lung <- read.csv(here("data/analysis_data/merged_data.csv"))
data_heart <- read.csv(here("data/analysis_data/merged_heart_data.csv"))
```

@@

# Plot the regression graph
ggplot(data, aes(x = provincial_average, y = total_deaths)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +  # Add linear regression line without confidence interval
  labs(x = "Provincial Average PM2.5", y = "Total Deaths") +
  theme_minimal()

@@

```{r}
#| label: fig-lung
#| fig-cap: Lung Related Causes Mortality Rates vs Average PM2.5 Quantity 
#| echo: false
#| warning: false

# Plot the regression graph for all lungs causes
ggplot(data_lung, aes(x = provincial_average, y = total_deaths)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +  # Add linear regression line without confidence interval
  labs(x = "Provincial Average", y = "Total Deaths") +
  theme_minimal()
```

```{r}
#| label: fig-heart
#| fig-cap: Heart Related Causes Mortality Rates vs Average PM2.5 Quantity 
#| echo: false
#| warning: false
# Plot the regression graph for all heart causes
ggplot(data, aes(x = provincial_average, y = total_deaths)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +  # Add linear regression line without confidence interval
  labs(x = "Provincial Average", y = "Total Deaths") +
  theme_minimal()

@@

Talk more about it.

@@

```{r}
#| label: fig-annual
#| fig-cap: Annual Trends of PM2.5 Concentrations in Alberta
#| echo: false
#| warning: false
#| message: false

# Assuming your data frame is named 'cleaned_avg_PM25_conc' and is already filtered and renamed.
# Convert factors to character to avoid issues in ggplot
data[] <- lapply(data, function(x) if(is.factor(x)) as.character(x) else x)
# Creating the plot with ggplot
p <- ggplot(data, aes(x = year)) +
  geom_line(aes(y = provincial_average, colour = "Provincial Average")) +
  geom_line(aes(y = x10th_percentile, colour = "10th Percentile")) +
  geom_line(aes(y = x90th_percentile, colour = "90th Percentile")) +
  labs(x = "Years", y = "PM2.5 Concentration (μg/m³)") +
  scale_colour_manual(values = c("Provincial Average" = "blue", "10th Percentile" = "green", "90th Percentile" = "red")) +
  theme_minimal()

# Display the plot
p
```

```{r}
#| label: fig-four
#| fig-cap: Mortality Rates vs. Air Quality Health Index
#| echo: false
#| warning: false
#| message: false

library(MASS)

air_quality_data <- air_quality_data[air_quality_data$air_quality_health_index != "10+",]
air_quality_data$air_quality_health_index <- as.numeric(air_quality_data$air_quality_health_index)

# Calculate the average Air Quality Health Index per year
average_aqi_by_year <- air_quality_data %>%
  group_by(year) %>%
  summarize(average_aqi = mean(air_quality_health_index))

combined_data <- merge(deaths_data, average_aqi_by_year, by = "year")

# Plot scatter plot of mortality rates vs. air quality health index
ggplot(combined_data, aes(x = average_aqi, y = total_deaths)) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_wrap(~cause) +
  labs(
       x = "Average Air Quality Health Index",
       y = "Total Deaths")
```

```{r}
#| label: fig-cause
#| fig-cap: Total Deaths by Year for Each Cause
#| echo: false
#| warning: false
#| message: false

# Summarize total deaths for each disease by year
total_deaths_by_year <- deaths_data %>%
  group_by(year, cause) %>%
  summarize(total_deaths = sum(total_deaths))

# Plot for mortality data
ggplot(total_deaths_by_year, aes(x = as.factor(year), y = total_deaths, group = cause, color = cause)) +
  geom_line() +
  labs(
       x = "Year",
       y = "Total Deaths") +
  theme(legend.position = "top") +
  theme(legend.text = element_text(size = 5)) +
  guides(color = guide_legend(direction = "vertical"))
```

```{r}
#| label: fig-average_air
#| fig-cap: Average Air Quality Health Index by Year
#| echo: false
#| warning: false
#| message: false
# Ensure the 'year' column is in numeric format for proper line plotting
average_aqi_by_year$year <- as.numeric(as.character(average_aqi_by_year$year))

# Plotting with ggplot2
ggplot(average_aqi_by_year, aes(x = year, y = average_aqi)) +
  geom_line() +  # for lines
  geom_point() + # for points
  labs(
       x = "Year",
       y = "Average Air Quality Index") +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 10)) # Optional: Adjust x-axis breaks
```
Talk way more about it.

# Model

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

\begin{align}
y_i | \mu_i, \phi &\sim \text{NegBin}(\mu_i, \phi) \\
\mu_i &= \exp(\alpha + \beta x_i) \\
\alpha &\sim \text{Normal}(0, 2.5) \\
\beta &\sim \text{Normal}(0, 2.5) \\
\phi &\sim \text{Exponential}(1)
\end{align}

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results

Our results are summarized in @tbl-modelresults.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

library(rstanarm)
library(here)
library(ggplot2)
library(modelsummary)

# Assuming you've already read your models as before
neg_binom_model <- readRDS(file = here::here("models/neg_binom_model.rds"))
stan_neg_binom_model <- readRDS(file = here::here("models/stan_neg_binom_model.rds"))
stan_poisson_model <-readRDS(file = here::here("models/stan_poisson_model.rds"))

heart_model <- readRDS(file = here::here("models/heart_model.rds"))
lung_model <-readRDS(file = here::here("models/lung_model.rds"))

first_model <-
  readRDS(file = here::here("models/first_model.rds"))

poisson_model <-
  readRDS(file = here::here("models/poisson_model.rds"))

# # Make predictions for heart and lung models
# heart_predictions <- predict(heart_model, newdata = data.frame(provincial_average = heart$provincial_average), se.fit = TRUE)
# lung_predictions <- predict(lung_model, newdata = data.frame(provincial_average = lung$provincial_average), se.fit = TRUE)
# # Create trend graphs
# # For example, let's plot predicted values against provincial average for heart data
# heart_pred_df <- data.frame(provincial_average = heart$provincial_average,
#                             total_deaths_pred = heart_predictions$fit,
#                             total_deaths_pred_upper = heart_predictions$fit + 1.96 * heart_predictions$se.fit,
#                             total_deaths_pred_lower = heart_predictions$fit - 1.96 * heart_predictions$se.fit)
# 
# # Plot
# ggplot(heart_pred_df, aes(x = provincial_average, y = total_deaths_pred)) +
#   geom_line() +
#   geom_ribbon(aes(ymin = total_deaths_pred_lower, ymax = total_deaths_pred_upper), alpha = 0.2) +
#   labs(x = "Provincial Average", y = "Predicted Total Deaths", title = "Predicted Total Deaths vs. Provincial Average (Heart Data)") +
#   theme_minimal()

# class(heart_model$provincial_average)
# class(lung_model$provincial_average)
# 
# # Create a range of values for the predictor variable
# provincial_average_range <- seq(min(heart$provincial_average, lung$provincial_average),
#                                 max(heart$provincial_average, lung$provincial_average),
#                                 length.out = 100)
# 
# # Make predictions for heart and lung models
# heart_predictions <- predict(heart_model, newdata = data.frame(provincial_average = provincial_average_range), se.fit = TRUE)
# lung_predictions <- predict(lung_model, newdata = data.frame(provincial_average = provincial_average_range), se.fit = TRUE)
# 
# # Create data frames for plotting
# heart_pred_df <- data.frame(provincial_average = provincial_average_range,
#                             total_deaths_pred = heart_predictions$fit,
#                             total_deaths_pred_upper = heart_predictions$fit + 1.96 * heart_predictions$se.fit,
#                             total_deaths_pred_lower = heart_predictions$fit - 1.96 * heart_predictions$se.fit)
# 
# lung_pred_df <- data.frame(provincial_average = provincial_average_range,
#                            total_deaths_pred = lung_predictions$fit,
#                            total_deaths_pred_upper = lung_predictions$fit + 1.96 * lung_predictions$se.fit,
#                            total_deaths_pred_lower = lung_predictions$fit - 1.96 * lung_predictions$se.fit)
# 
# # Plot predicted deaths for heart and lung models
# ggplot() +
#   geom_line(data = heart_pred_df, aes(x = provincial_average, y = total_deaths_pred, color = "Heart Data")) +
#   geom_ribbon(data = heart_pred_df, aes(x = provincial_average, ymin = total_deaths_pred_lower, ymax = total_deaths_pred_upper, fill = "Heart Data"), alpha = 0.2) +
#   geom_line(data = lung_pred_df, aes(x = provincial_average, y = total_deaths_pred, color = "Lung Data")) +
#   geom_ribbon(data = lung_pred_df, aes(x = provincial_average, ymin = total_deaths_pred_lower, ymax = total_deaths_pred_upper, fill = "Lung Data"), alpha = 0.2) +
#   labs(x = "Provincial Average", y = "Predicted Total Deaths", title = "Predicted Total Deaths due to Air Pollution") +
#   scale_color_manual(values = c("Heart Data" = "blue", "Lung Data" = "red")) +
#   scale_fill_manual(values = c("Heart Data" = "blue", "Lung Data" = "red"), guide = FALSE) +
#   theme_minimal()

@@

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelcauses
#| tbl-cap: "Different causes of mortality and their death counts"
#| warning: false
library(modelsummary)

if (!requireNamespace("modelsummary", quietly = TRUE)) {
  install.packages("modelsummary")
}
library(modelsummary)

if (!requireNamespace("broom.mixed", quietly = TRUE)) {
  install.packages("broom.mixed")
}

library(broom.mixed)

coef_short_names <- 
  c("causeAll other forms of chronic ischemic heart disease" = "Ischemic Heart Disease",
    "causeMalignant neoplasms of trachea, bronchus and lung" = "Trachea/Bronchus/Lung Cancer",
    "Acute myocardial infarction" = "Heart Attack",
    "causeOther chronic obstructive pulmonary disease" = "COPD"
  )

# Summarize the models
modelsummary(
  list(
    "Poisson" = poisson_model,
    "Negative binomial" = first_model
  ),
  coef_map = coef_short_names,
  estimate_function = function(model) {
    broom.mixed::tidy(model, parameters = "all", conf.int = TRUE)
  }
)

@@

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelheart
#| tbl-cap: "Heart related causes death count and the air quality"
#| warning: false

modelsummary::modelsummary(
  list(
    "heart causes model" = heart_model
  ),
  #statistic = "mad",
  fmt = 2
)

@@

```{r}
#| echo: false
#| eval: true
#| label: tbl-modellung
#| tbl-cap: "Lung related causes death count and the air quality"
#| warning: false
modelsummary::modelsummary(
  list(
    "Lung Causes model" = lung_model
  ),
  #statistic = "mad",
  fmt = 2
)

```
# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\appendix

# Appendix {-}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckfirst
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

pp_check(first_model) +
  theme(legend.position = "bottom")

pp_check(poisson_model) +
  theme(legend.position = "bottom")
```

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckheartlung
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

pp_check(heart_model) +
  theme(legend.position = "bottom")

pp_check(lung_model) +
  theme(legend.position = "bottom")

@@

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

plot(first_model, "trace")

plot(first_model, "rhat")

plot(stan_neg_binom_model, "trace")

plot(stan_neg_binom_model, "rhat")
```

>>>>

breach_data <- read.csv(here("data/analysis_data/breach_data.csv"))
```

@@

breach_data %>%
  count(year) %>%
  ggplot(aes(x = year, y = n)) +
  geom_line() +
  labs(title = "Cyberattacks Over Time", x = "Year", y = "Number of Attacks")

``` 
```{r}
breach_data %>%
  count(sector) %>%
  ggplot(aes(x = reorder(sector, n), y = n)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  labs(title = "Cyberattacks by Sector", x = "Sector", y = "Number of Attacks")
```

```{r}
breach_data %>%
  count(country) %>%
  ggplot(aes(x = reorder(country, n), y = n)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  labs(title = "Cyberattacks by Country", x = "Country", y = "Number of Attacks")

@@

```{r}
breach_data %>%
  count(country) %>%
  ggplot(aes(x = reorder(country, n), y = n, fill = country)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  labs(title = "Cyberattacks by Country", x = "Country", y = "Number of Attacks") +
  theme_minimal() +
  theme(legend.position = "none") # Hides the legend

@@

```{r}
breach_data %>%
  count(attack_type) %>%
  ggplot(aes(x = "", y = n, fill = attack_type)) +
  geom_bar(width = 1, stat = "identity") +
  coord_polar("y") +
  labs(title = "Distribution of Attack Types")

```
Talk more about it.

@@

```{r}
breach_data %>%
  ggplot(aes(x = as.factor(year), y = impact_on_data)) +
  geom_boxplot() +
  labs(title = "Impact on Data by Year", x = "Year", y = "Impact on Data")

```
```{r}
breach_data %>%
  count(year, attack_type) %>%
  ggplot(aes(x = year, y = n, fill = attack_type)) +
  geom_area(position = 'stack') +
  labs(title = "Attack Types Over Years", x = "Year", y = "Count")

@@

```{r}
library(ggplot2)
library(dplyr)

# Assuming 'overall_nature_of_attack' is your categorical variable indicating the nature/type of attack
# and 'number_of_users_affected' is a numeric variable indicating the number of users impacted

# First, let's calculate summary statistics for each nature of attack

breach_data %>%
  count(year, sector) %>%
  ggplot(aes(x = year, y = sector, fill = n)) +
  geom_tile() +
  scale_fill_gradient(low = "lightblue", high = "red") +
  labs(title = "Heatmap of Cyberattacks per Year and Sector",
       x = "Year",
       y = "Sector")

```
```{r}
breach_data %>%
  count(year, overall_nature_of_attack) %>%
  ggplot(aes(x = as.factor(year), y = n, fill = overall_nature_of_attack)) +
  geom_bar(stat = "identity") +
  labs(title = "Stacked Bar Chart of Attack Types by Year",
       x = "Year",
       y = "Number of Attacks")

```
```{r}
library(ggplot2)

ggplot(breach_data, aes(x = as.factor(year), y = number_of_users_affected)) +
  geom_boxplot() +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Boxplot of Number of Users Affected Across Years",
       x = "Year",
       y = "Number of Users Affected")
```

```{r}

breach_data %>%
  filter(year == 2013) %>%
  arrange(desc(number_of_users_affected)) %>%
  head()  # This shows the top entries for 2013

breach_data %>%
  filter(year == 2014) %>%
  arrange(desc(number_of_users_affected)) %>%
  head()  # This shows the top entries for 2013

# Install and load the scales package
if (!requireNamespace("scales", quietly = TRUE)) {
  install.packages("scales")
}
library(scales)

@@

```{r}
breach_data %>%
  count(country) %>%
  mutate(percentage = n / sum(n) * 100) %>%
  ggplot(aes(x = reorder(country, percentage), y = percentage)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  labs(title = "Percentage of Total Cyberattacks by Country",
       x = "Country",
       y = "Percentage of Attacks")

@@

```{r}

breach_data %>%
  mutate(impact_score = case_when(
    impact_on_data == "Low" ~ 1,
    impact_on_data == "Medium" ~ 2,
    impact_on_data == "High" ~ 3,
    TRUE ~ NA_real_
  )) %>%
  group_by(organisation_size) %>%
  summarize(average_impact = mean(impact_score, na.rm = TRUE)) %>%
  ggplot(aes(x = organisation_size, y = average_impact, fill = organisation_size)) +
  geom_col() +
  labs(title = "Average Impact Severity by Organization Size",
       x = "Organisation Size",
       y = "Average Impact Score") +
  theme_minimal()

@@

```{r}
ggplot(breach_data, aes(x = number_of_users_affected)) +
  geom_density(fill = "blue", alpha = 0.5) +
  labs(title = "Density Plot of Number of Users Affected by Cyberattacks",
       x = "Number of Users Affected",
       y = "Density")

```{r}
ggplot(breach_data, aes(x = organisation_size, fill = organisation_size)) +
  geom_bar() +
  labs(title = "Number of Cyberattacks by Organization Size",
       x = "Organization Size",
       y = "Number of Cyberattacks") +
  theme_minimal()

```
```{r}
library(modelsummary)

logistic_model <- readRDS(file = here::here("models/restructuring_model.rds"))

modelsummary(list("Logistic Regression" = logistic_model))

```
```{r}
breach_data <- breach_data %>% mutate(row_id = row_number())

# Adjust factors in your data to match the model's training data
breach_data <- breach_data %>%
  mutate(country = factor(country, levels = levels(logistic_model$model$country)))

# Generate predictions
breach_predictions <- predict(logistic_model, newdata = breach_data, type = "response")

# Combine the predictions with the original data
breach_data <- breach_data %>% mutate(predicted_prob = breach_predictions)

# Scatter plot with jitter
ggplot(breach_data, aes(x = organisation_size, y = predicted_prob)) +
  geom_jitter(alpha = 0.3) +
  labs(x = "Organisation Size", y = "Predicted Probability of Restructuring")
